---
title: "Handout- Probability"
subtitle: Jacob Thoma
output:
  html_document: default
  #word_document: default
---
## On your own

### Comparing Kobe Bryant to the Independent Shooter

Using `calc_streak`, compute the streak lengths of `sim_basket`.

1.   Describe the distribution of streak lengths. What is the typical streak 
    length for this simulated independent shooter with a 45% shooting percentage?
    How long is the player's longest streak of baskets in 133 shots?

     ***Answer:*** The simulated player had a typical streak length of zero, meaning that 
     their shots were not followed by a shot with a similar result (Hit or Miss). The
     longer the streaks became the less there were. The players longest streak was 7.  
     ***NOTE:*** When HMTL is knit and everything is re-run the numbers on the table do not typically match 
     the answers I gave. 
    
```{r}
# Load Data & calc_streak func
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")

# Prepare data
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 133, replace = TRUE)

# Calculate and examine simulated player streak 
sim_streak <- calc_streak(sim_basket)
barplot(table(sim_streak))
```

2.   If you were to run the simulation of the independent shooter a second time, 
    how would you expect its streak distribution to compare to the distribution 
    from the question above? Exactly the same? Somewhat similar? Totally 
    different? Explain your reasoning.

     ***Answer:***  
     ***Before Run:*** I would expectthe distribution to be very simailar, at least in the patterns it 
     revealed, but I would expect it to be slightly different perhaps with a shorter or longer 'longest 
     streak' length.  
     ***After Run:*** I was exactly right, the distribution is similar, but this time the longest streak
     was 8!  
     ***NOTE:*** When HMTL is knit and everything is re-run the numbers on the table do not typically match 
     the answers I gave. 
    
```{r}
# Load Data & calc_streak func
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")

# Prepare data
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 133, replace = TRUE)

# Calculate and examine simulated player streak 
sim_streak <- calc_streak(sim_basket)
barplot(table(sim_streak))
```

3.   How does Kobe Bryant's distribution of streak lengths compare to the 
    distribution of streak lengths for the simulated shooter? Using this 
    comparison, do you have evidence that the hot hand model fits Kobe's 
    shooting patterns? Explain.
    
     ***Answer:*** Although I am dismayed to say it the distribution of Kobe's
     shots and the simulated player's shots are very similar, pointing away from 
     from the hot hand model for Kobe's shooting pattern. If Kobe's shot results 
     were a result of a hot hands model we would expect to see more actual streak
     lengths that were greater than 0. Additionaly we would likely expect the 
     maxiumum streak length to be longer than that of the simulated player. 
    
```{r}
# Load Data & calc_streak func
download.file("http://www.openintro.org/stat/data/kobe.RData", destfile = "kobe.RData")
load("kobe.RData")

# Calculate and plot any Kobe streaks
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))

# Prepare sim data data
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 133, replace = TRUE)

# Calculate and examine simulated player streak 
sim_streak <- calc_streak(sim_basket)
barplot(table(sim_streak))
```    

<div id="license">
This is a product of OpenIntro that is released under a 
[Creative Commons Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0). 
This lab was adapted for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel
from a lab written by Mark Hansen of UCLA Statistics.
</div>