---
title: "Handout- Sampling Distributions"
subtitle: Jacob Thoma
output:
  html_document: default
  #word_document: default
---

## On your own

So far, we have only focused on estimating the mean living area in homes in 
Ames.  Now you'll try to estimate the mean home price.

1.   Take a random sample of size 50 from `price`. Using this sample, what is your best point estimate of the population mean?

     ***Answer:  
     Taking the mean of the sample is our best point estimate of the population mean, 
     in this case it is 169,076.7***
    
```{r}
download.file("http://www.openintro.org/stat/data/ames.RData", destfile = "ames.RData")
load("ames.RData")

area <- ames$Gr.Liv.Area
price <- ames$SalePrice

samp2 <- sample(price, 50)
mean(samp2)
```
     
2.   Since you have access to the population, simulate the sampling 
    distribution for $\bar{x}_{price}$ by taking 5000 samples from the 
    population of size 50 and computing 5000 sample means.  Store these means 
    in a vector called `sample_means50`. Plot the data, then describe the 
    shape of this sampling distribution. Based on this sampling distribution, 
    what would you guess the mean home price of the population to be? Finally, 
    calculate and report the population mean.

     ***Answer:  
     The histogram of the sample means is very symetrical. Based on the histogram alone 
     I'd guess the mean population home proce would be very close to 180k.  
     After calculating the actual population mean I got 180,796.1 which is very close to 
     my estimate based off the samples.
     ***
    
```{r}
sample_means50 <- rep(NA, 5000)

for(i in 1:5000){
  samp <- sample(price, 50)
  sample_means50[i] <- mean(samp)
}

hist(sample_means50, breaks = 25)

mean(price)
```
     
3.   Change your sample size from 50 to 150, then compute the sampling 
    distribution using the same method as above, and store these means in a 
    new vector called `sample_means150`. Describe the shape of this sampling 
    distribution, and compare it to the sampling distribution for a sample 
    size of 50.  Based on this sampling distribution, what would you guess to 
    be the mean sale price of homes in Ames?

     ***Answer:  
     Again the sampling distribution is fairly symetrical but the peak appears to 
     have moved higher than it was when we only had 50 samples. Based on this
     data distribution I would predict the mean to be around 181k (which is 
     closer to the known mean)***
    
```{r}
sample_means150 <- rep(NA, 5000)

for(i in 1:5000){
  samp <- sample(price, 150)
  sample_means150[i] <- mean(samp)
}

hist(sample_means150, breaks = 25)
     
```
     
4.   Of the sampling distributions from 2 and 3, which has a smaller spread?  If
    we're concerned with making estimates that are more often close to the 
    true value, would we prefer a distribution with a large or small spread?

     ***Answer:  
     The spread is smaller on the distribution where our sample was 150 which is
     preferable when we are trying to hone in on making estimates closse to the 
     true value. ***
    
```{r}

     
```
     
<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was written for OpenIntro by Andrew Bray and Mine &Ccedil;etinkaya-Rundel.
</div>