---
title: "Handout-Introduction to Data"
subtitle: Jacob Thoma 
output:
  html_document: default
  word_document: default
---
## On Your Own

1.   Make a scatterplot of weight versus desired weight. Describe the 
    relationship between these two variables.

     ***Answer:*** It seems that as the actual weight increases, so does the desired
     weight of the individual, indicating that most people are looking to only
     gain or loose just a few punds, with far less people looking for drastic 
     weight changes. There also appears to be several outliers.
    
```{r}
source("http://www.openintro.org/stat/data/cdc.R")
plot(x = cdc$wtdesire, y = cdc$weight, type = 'p', main = "Actual Weight vs. Desired Weight", 
     xlab = 'Desired Weight', ylab = 'Actual Weight')
```

2.   Let's consider a new variable: the difference between desired weight 
    (`wtdesire`) and current weight (`weight`). Create this new variable by 
    subtracting the two columns in the data frame and assigning them to a new 
    object called `wdiff`.
    
```{r}
cdc$wdiff = cdc$wtdesire - cdc$weight
```

3.   What type of data is `wdiff`? If an observation `wdiff` is 0, what does 
    this mean about the person's weight and desired weight. What if `wdiff` is 
    positive or negative?

     ***Answer:*** The 'wdiff' field contains numerical continuous data. If 'wdiff'
     os 0 it means that the individual didn't want to loose or gain any weight 
     (their actual and desired weight are the same). If 'wdiff' is positive it 
     means the individual wanted to gain weight, and if it is negative then the 
     individual wanted to loose weight.
     
4.   Describe the distribution of `wdiff` in terms of its center, shape, and 
    spread, including any plots you use. What does this tell us about how people 
    feel about their current weight?

      ***Answer:***  
      ***'wdiff' Data Summary:***  
      Min. 1st Qu.  Median    Mean 3rd Qu.    Max.  
      -300.00  -21.00  -10.00  -14.59    0.00  500.00 
      
      Looking at the data summary alone we can see that the median is -10,
      meaning that most poeple wanted to loose 10 pounds. Interestly enough we 
      can also see that the 3rd Qu. is 0, meaning that more of the data fell below
      0 than above. This translates to observing that more people wanted too loose
      weight than gain it. The average amount of weight people wanted to change by 
      was -14.59. 
      From the histogram we can see the data is unimodal and skewed to the left, 
      again indicating more poeple wanted to loose weight than to gain it. Because
      of the unimodality we can also support the claim that a majority of people 
      also only wanted to change their weight a little bit instead of drastic
      changes. 
    
```{r}
summary(cdc$wdiff)
hist(cdc$wdiff, col = blues9, main = 'Desired Weight Change', 
     xlab = 'Weight Change (lbs)' ) 
```
     
5.   Using numerical summaries and a side-by-side box plot, determine if men tend
    to view their weight differently than women.

     ***Answer:***  
     ***Male Weight Difference Data Summary:***  
     Min. 1st Qu.  Median    Mean 3rd Qu.    Max.   
    -300.00  -20.00   -5.00  -10.71    0.00  500.00   
     ***Female Weight Difference Data Summary:***  
     Min. 1st Qu.  Median    Mean 3rd Qu.    Max.  
    -300.00  -27.00  -10.00  -18.15    0.00   83.00  
    Looking at the data summary we can see that the 1st quarters (low end) are much
    lower with females than males. The same is true for the median and the mean. This
    indicates that females typically want to loose more weight than males. However it is
    worth noting that both male and female's 3rd quarter was at 0, following the larger
    trend we observed that more people want to loose weight than gain it. 
    When making the box plots, I excluded the possible outliers to make the data easier to
    see. We can see the distance between the #rd and 1st qu. on the female box plot is 
    greater than that of the male box plot, indicating that the amount of weight females
    wanted to loose had more variance than males, who typically wanted to loose less weight. 

    
```{r}
mdata <- subset(cdc, cdc$gender == 'm')
fdata <- subset(cdc, cdc$gender == 'f')

summary(mdata$wdiff)
summary(fdata$wdiff)

# Drew Box plot and excluded outliers to see data beter
boxplot(cdc$wdiff ~ cdc$gender, outline=FALSE, 
        main = 'Desired Weight Change: Male vs. Female', 
        ylab = 'Weight Chnage', xlab = 'Gender')
```
     
6.   Now it's time to get creative. Find the mean and standard deviation of 
    `weight` and determine what proportion of the weights are within one 
    standard deviation of the mean.

     ***Answer:***  
     70.76% off all the data fall within 1 standard deviation of the mean. This fits with what
     we know about a perfect distrubution containing 68% of all data within 1 standard 
     deviation of the mean. 
    
```{r}
alldata <- cdc$weight
total <- length(alldata)

pm1std <- length(alldata[abs(alldata - mean(alldata)) < sd(alldata)])

ans <- pm1std / total 

ans*100
```
