---
title: "Handout-Introduction to R and RStudio"
subtitle: Jacob Thoma
output:
  html_document: default
  word_document: default
---
## On Your Own

In the previous few pages, you recreated some of the displays and preliminary 
analysis of Arbuthnot's baptism data. Your assignment involves repeating these 
steps, but for present day birth records in the United States.  Load up the 
present day data with the following command.

```{r load-present-data, eval=FALSE}
source("http://www.openintro.org/stat/data/present.R")
```

The data are stored in a data frame called `present`.

1.   What years are included in this data set? What are the dimensions of the 
    data frame and what are the variable or column names?

     ***Answer: The data includes years 1940 - 2002. The dimensions of th data frame 
     are 63 rows x 3 columns. The column names are 'year', 'boys', and 'girls'. ***
    
```{r}
source("http://www.openintro.org/stat/data/present.R")

range(present$year)
dim(present)
names(present)
```
    
2.   How do these counts compare to Arbuthnot's? Are they on a similar scale?

     ***Answer: The present day data and Arbuthnot's data are comparable, they have
     the same 3 variables but Arbuthnot's data has records for 19 more years.***

```{r}
source("http://www.openintro.org/stat/data/present.R")
source("http://www.openintro.org/stat/data/arbuthnot.R")
dim(present)
dim(arbuthnot)
range(arbuthnot$year)
```

3.   Make a plot that displays the boy-to-girl ratio for every year in the data 
    set. What do you see? Does Arbuthnot's observation about boys being born in
    greater proportion than girls hold up in the U.S.? Include the plot in your 
    response.

     ***Answer: Arbuthnot's observation about boys being born in greater proportion 
     than girls seems to hold up in the U.S. even several centuries later. ***

```{r}
source("http://www.openintro.org/stat/data/present.R")
plot(x = present$boys, y = present$girls, type = 'p', main = "Boys and Girls Born by Year (1940-2002)", 
     xlab = 'Boys', ylab = 'Girls')
```

4.   In what year did we see the most total number of births in the U.S.? You can
    refer to the help files or the R reference card 
    [http://cran.r-project.org/doc/contrib/Short-refcard.pdf](http://cran.r-project.org/doc/contrib/Short-refcard.pdf)
    to find helpful commands.

     ***Answer:  There were the most births in 1961. There were 4,268,326 births. ***

```{r}
source("http://www.openintro.org/stat/data/present.R")
present$Totals = present$boys + present$girls
maxBirths <- max(present$Totals)
present$year[present$Totals == maxBirths]
```

These data come from a report by the Centers for Disease Control 
[http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf](http://www.cdc.gov/nchs/data/nvsr/nvsr53/nvsr53_20.pdf). 
Check it out if you would like to read more about an analysis of sex ratios at 
birth in the United States.

That was a short introduction to R and RStudio, but we will provide you with more
functions and a more complete sense of the language as the course progresses. 
Feel free to browse around the websites for [R](http://www.r-project.org) and 
[RStudio](http://rstudio.org) if you're interested in learning more, or find 
more labs for practice at http://openintro.org.

